The Ranked Nearest Neighbour Interchange space

### 2019-06-26 paper.tex
add figure RNNI moves definition
edits on proof: caterpillar convex

\todo{figure: rank + $\nni$ move}

@@

    Induction basis: $l = \Delta(\rnniu) = \frac{(n-1)(n-2)}{2}$\\
    With Lemma~\ref{lemma:caterpillar_dist=diameter} one can see that for each pair of caterpillar trees $T$, $R$ with caterpillar distance $d_c(T,R) = \Delta(\rnniu)$, it is $d(T,R) = d_c(T,R)$, which means that there is a shortest path between $T$ and $R$ that is a caterpillar path.

    Induction step: $l+1 \to l$\\
    Induction hypothesis: For each pair of caterpillar trees with caterpillar distance $l+1$ there is a shortest path between these trees that only consists of caterpillar trees, which means that the distance between these trees is $l+1$.
    Let $T$ and $R$ be caterpillar trees with caterpillar distance $l$.
    Since $l < \Delta(\rnniu) = \frac{(n-1)(n-2)}{2}$, there is one loop $k$ in the Bubble Sort algorithm from $R$ to $T$ where there are less than $n-2-k$ swaps of neighboured taxa.
    This means that there is a taxon $a$ that has parent with rank $n-2-k$ in $T$ and that does not need to move all the way up starting at the cherry in loop $k$.
    Therefore, there must be a taxon $b$ which has parent with rank one less than the parent of $a$ in the tree at the beginning of loop number $k$.
    If we assume that loop $k$ is the first loop where a single taxon is not moved by $n-k-2$ $\rnni$ moves, it follows that $a$ and $b$ must be neighboured taxa in $R$.
    Let $R'$ be the caterpillar tree resulting from $R$ by exchanging $a$ and $b$.
    It follows that $R$ has caterpillar distance $d_c(T,R') = d_c(T,R) + 1 = l+1$ to $T$ and $d(R,R') = 1$.
    We can apply the induction hypothesis on $T$ and $R'$ and know that $d(T,R') = l+1$.
    If it was $d(T,R) < l$, it would hold $d(T,R') \leq d(T,R) + d(R,R') < l + 1$ which contradicts the induction hypothesis.
    We can conclude that $d(T,R) = l$.

@@

>>>>

\begin{figure}[H]
	\centering
	\includegraphics[width=\textwidth]{RNNI}
	\caption{Four trees connected by edges in the $\rnni$ graph. A rank move of $mrca(\{1,2\})$ and $mrca(\{3,4\})$ and all $\nni$ moves on the edge highlighted in red are illustrated}
	\label{fig:NNI_vs_RNNI}
\end{figure}

\todo{label of taxa in figure shouldn't be numbers}

@@

    Induction basis: $l = \frac{(n-1)(n-2)}{2}$\\
    With Lemma~\ref{lemma:caterpillar_dist=diameter} one can see that for each pair of caterpillar trees $T$, $R$ with caterpillar distance $d_c(T,R) = \frac{(n-1)(n-2)}{2}$, it is $d(T,R) = d_c(T,R)$, which means that there is a shortest path between $T$ and $R$ that is a caterpillar path.

    Induction step: $l+1 \to l$\\
    Induction hypothesis: For each pair of caterpillar trees with caterpillar distance $l+1$ there is a shortest path between these trees that only consists of caterpillar trees, which means that the distance between these trees is $l+1$.
    Let $T$ and $R$ be caterpillar trees with caterpillar distance $l$.
    Since $l < \frac{(n-1)(n-2)}{2}$, there is one loop $k$ in the caterpillar path from $R$ to $T$ that is given by the Bubble Sort algorithm \todo{name} where less than $n-2-k$ pairs taxa swap positions.
    It follows that there is a taxon $a$ that has parent with rank $n-2-k$ in $T$ and does not need to move all the way up starting at the cherry in loop $k$.
    Let $b$ be the taxon whose parent has rank one less than the parent of $a$ in the tree at the beginning of loop number $k$.
    We can assume without loss of generality that loop $k$ is the first loop in which the number of swaps is not maximal.
    It follows that $a$ and $b$ must be neighboured taxa in $R$.
    Let us no consider caterpillar tree $R'$ resulting from $R$ by exchanging taxa $a$ and $b$.
    It is clear that $R'$ has caterpillar distance $d_c(T,R') = d_c(T,R) + 1 = l+1$ to $T$ and moreover, $d(R,R') = 1$.
    Hence we can apply the induction hypothesis on $T$ and $R'$ and see that $d(T,R') = l+1$.
    If it were $d(T,R) < l$, we would have $d(T,R') \leq d(T,R) + d(R,R') < l + 1$ which contradicts the induction hypothesis.
    We can conclude $d(T,R) = l$, which completes the proof.
\end{proof}

@@

\todo{This is already an implication of the results above. Shall we still keep a subsection for Diameter? In case we find something about the radius, this could be added here.}
